## Coursera Reproducible Research Project 2
The following code will take storm data from the national weather service and determine
the 10 most devestating type of weather events in terms of fatalities, injuries and economic impact.

# Data Processing
# Processing data for injuries and fatalities

```{r}
library(ggplot2)
zipfile <- bzfile("repdata%2Fdata%2FStormData.csv.bz2")
df1 <- read.csv(zipfile, header = TRUE)

## data for number of fatalities by natural eventype
fatalitiesDF <- aggregate(df1$FATALITIES, by=list(evtype1 = df1$EVTYPE), FUN = sum)
fatalitiesDF <- fatalitiesDF[order(fatalitiesDF$x, decreasing = TRUE), ]
fatalitiesDF <- head(fatalitiesDF, 10)

## data for number of injuries by natural eventype
injuriesDF <- aggregate(df1$INJURIES, by=list(evtype2 = df1$EVTYPE), FUN = sum)
injuriesDF <- injuriesDF[order(injuriesDF$x, decreasing = TRUE), ]
injuriesDF <- head(injuriesDF, 10)
```

# processing economic impact by event type
```{r}
knitr::opts_chunk$set(cache=TRUE)

df2 <- df1
df2$PROPDMGEXP <- as.character(df2$PROPDMGEXP)
df2$CROPDMGEXP <- as.character(df2$CROPDMGEXP)

for (i in 1:length(df2$PROPDMGEXP)){
        if (df2$PROPDMGEXP[i] == "H") {
                df2$PROPDMG[i] <- df2$PROPDMG[i] * 100
        } else if (df2$PROPDMGEXP[i] == "K") {
                df2$PROPDMG[i] <- df2$PROPDMG[i] * 1000        
        } else if (df2$PROPDMGEXP[i] == "M") {
                df2$PROPDMG[i] <- df2$PROPDMG[i] * 1000000
        } else if (df2$PROPDMGEXP[i] == "B") {
                df2$PROPDMG[i] <- df2$PROPDMG[i] * 1000000000
        } else {
                df2$PROPDMG[i] <- df2$PROPDMG[i]
        }
}

for (i in 1:length(df2$CROPDMGEXP)){
        if (df2$CROPDMGEXP[i] == "H") {
                df2$CROPDMG[i] <- df2$CROPDMG[i] * 100
        } else if (df2$CROPDMGEXP[i] == "K") {
                df2$CROPDMG[i] <- df2$CROPDMG[i] * 1000        
        } else if (df2$CROPDMGEXP[i] == "M") {
                df2$CROPDMG[i] <- df2$CROPDMG[i] * 1000000
        } else if (df2$CROPDMGEXP[i] == "B") {
                df2$CROPDMG[i] <- df2$CROPDMG[i] * 1000000000
        } else {
                df2$CROPDMG[i] <- df2$CROPDMG[i]
        }
}

df2$sumCost <- with(df2, PROPDMG + CROPDMG)

costDF <- aggregate(df2$sumCost, by=list(evtype = df2$EVTYPE), FUN = sum)
costDF <- costDF[order(costDF$x, decreasing = TRUE), ]
costDF <- head(costDF, 10)

```
## Results
# Number of fatalities
```{r fig.align = 'center', fig.height = 4}

plot1 <- ggplot(fatalitiesDF, aes(reorder(evtype1,x), x)) ## reorder w/ respect to x
plot1 + geom_bar(stat = "identity", fill = "coral1", width = .9) +
        coord_flip() +
        xlab("Event Type") +
        ylab("Number of Fatalities") +
        ggtitle("Fatalities per Event Type")         

```


# Number of injuries
```{r fig.align = 'center', fig.height = 4}

plot2 <- ggplot(injuriesDF, aes(reorder(evtype2, x), x))
plot2 + geom_bar(stat = "identity", fill = "coral1", width = .9) +
        coord_flip() +
        xlab("Event Type") +
        ylab("Number of Injuries") +
        ggtitle("Injuries per Event Type")         

```
Tornados seem to cause the greatest number of injuries by far compared
to any other type of event

# Economic impact of storm
```{r fig.align = 'center', fig.height = 4}

plot3 <- ggplot(costDF, aes(reorder(evtype, x), x))
plot3 + geom_bar(stat = "identity", fill = "coral1", width = .9) +
        coord_flip() +
        xlab("Event Type") +
        ylab("Economic impact") +
        ggtitle("Economic impact per Event Type")         
```
According to the data Floods seem to cause the greatest economic impact


